Easy as Pi: A Network Coding Raspberry Pi Testbed

### 2016-06-06 introduction.tex
Second pass on the introduction

At the beginning, there was darkness and then... bang! \ac{NC}
\cite{ahlswede2000network} appeared to save us from the evilness
of routing.

General introduction. Introduction to topic addressed in the journal.
Review of the State of the Art. Specify why our approach has benefits
and which are they. Indicate contributions.

Our work is organized as follows: Section \ref{sec:testbed} introduces
the testbed system, its setup, connectivity, files configuration and
compilation of the Kodo library. Section \ref{sec:schemes} defines the
coding schemes employed in our study. Later, in Section \ref{sec:metrics}
we describe the considered metrics for performance comparison of the
codes deployed in the \ac{Raspi}. In Section \ref{sec:measurements}, we show
the measurements in the \ac{Raspi} of the mentioned metrics. Final
conclusions and future work are reviewed in Section \ref{sec:conclusions}.

>>>>

%!TEX root = raspi_journal.tex
% At the beginning, there was darkness and then... bang! \ac{NC}
% \cite{ahlswede2000network} appeared to save us from the evilness
% of routing.
%
% General introduction. Introduction to topic addressed in the journal.
% Review of the State of the Art. Specify why our approach has benefits
% and which are they. Indicate contributions.

The emergence of powerful and inexpensive single-board computers
running full versions of standard operating systems (OS) in the last
years opens exciting possibilities for researchers. This not only
allows to seamlessly develop implementations that are compatible with
higher end devices and allows the reuse stable software from the Open
Source community. It also enables the deployment and testing of large
scale (hundreds of devices or more) distributed systems at a fraction
of the cost necessary in previous years. This paper focuses on
providing the critical steps and mechanisms followed to deploy, setup,
and maintain a large-scale testbed based on Raspberry Pi devices of
different models. We also complement this with an in-depth evaluation
of the Raspberry Pi as a platform to enable a new paradigm in
networking and storage known as network coding, which relies on
processing at intermediate nodes to achieve higher throughput, delay
and reliability performance than state-of-the-art routing techniques.

% INTRODUCTION TO NETWORK CODING
Introduced by Ahlswede et al~\cite{ahlswede2000network}, network
coding constitutes paradigm shift in the way how researchers and
industry understand and operate networks, by changing the role of
intermediate relays in the process of transmission of information.
Relays are no longer limited to storing and forwarding data, but also
take part in the coding process, through a process called recoding,
where the relay generates new linear combinations of incoming coded
packets without previously decoding the data. Network coding allows
the increase of throughput, reliability, security and delay
performance of the networks.

Compared to other broadly used coding schemes, e.g., Gallager's LDPC
codes~\cite{gallager1962low}, Reed-Solomon
codes~\cite{reed1960polynomial}, network coding is a technology that
has been implemented in real systems since the early years of its
conception. In 2006, Katabi et al~\cite{katabi2006practical} published
the results of the performance of their protocol COPE when implemented
in a real wireless system, which relied on minimalistic coding to
provide interesting gains. In 2008, Pedersen et
al~\cite{pedersen2008implementation} used commercially available
Symbian OS enabled mobile phones to implement network coding in a
\ac{D2D} cooperation application. In 2011, researchers at Aalborg
University developed KODO~\cite{kodo2011pedersen}, a C++ network
coding library with open source code for researchers intended to make
network coding implementations easily available for both, the research
community and commercial entities.
CATWOMAN~\cite{hundeboll2012catwoman} is a protocol implemented on top
of the B.A.T.M.A.N. protocol~\cite{johnson2008simple} for IEEE 802.11
multi-hop meshed networks using some of the intuition from COPE, but
deploying it in real systems. It is currently available as open source
code in the Linux kernel. Many other implementations have been tested
on real world systems~
\cite{pahlevani2013playncool,katti2008xors,krigslund2013core,paramanathan2013leanandmean}.

Despite this successful deployment in real systems, many of these
protocols and contributions have been implemented in separate testbeds
and the experiences are hard to reproduce.  Given the need for
reproducibility of past and future results among researchers and the
need of low-cost and easy-to-deploy testbeds, this paper contribution
is provide detailed instructions on how to set up a low-cost,
heterogeneous and scalable testbed consisting of Raspberry Pis
(Raspis), and by providing detailed measurements of goodput and energy
consumption of such testbed when performing network coding operations
with different codecs (such as full RLNC, multicore-enabled RLNC,
sparse RLNC and tunable sparse RLNC). The heterogeneity of the testbed
comes from the fact that old, single core, Raspberry Pi 1 model B are
integrated seamlessly in the described testbed with newer, multicore
versions, like the Raspberry Pi 2 model B.

Our work is organized as follows. Section~\ref{sec:testbed} introduces the
testbed system, its setup, connectivity, files configuration and compilation of
the Kodo library. Section~\ref{sec:schemes} defines the coding schemes employed
in our study. Later, in Section~\ref{sec:metrics} we describe the considered
metrics for performance comparison of the codes deployed in the \ac{Raspi}. In
Section~\ref{sec:measurements}, we show the measurements in the \ac{Raspi} of
the mentioned metrics. Final conclusions and future work are reviewed in
Section~\ref{sec:conclusions}.
